RIScatter: Unifying\\Backscatter Communications, Symbiotic Radio, and Reconfigurable Intelligent Surface

### 2022-08-30 main.tex
:pencil: Update abstract

	\gls{bc} switches load impedance to manipulate the magnitude, phase, and/or frequency responses and modulate information onto the scattered signal.
	\gls{ris} programs antenna/metamaterial array to adjust the scattering phase shifts and control signal strength in specific directions.

@@

	\gls{sr} introduces a passive scatterer to ride secondary information over the primary signal generated by an active transmitter, and employs a co-located receiver to sequentially decode both links.

@@

	RIScatter is a new paradigm for future wireless networks and consists of multiple passive scattering nodes, whose reflection states can be controlled to partially engineer the wireless channel and partially modulating information onto the scattered wave.
	This contrasts with \gls{bc}/\gls{sr} (resp. \gls{ris}) where states are exclusively a function on information symbols (resp. \gls{csi}).
	The key principle in RIScatter is to render the probability distribution of reflection states as a joint function of the \gls{csi} and input information source, allowing universal hardware design and heterogeneous traffic control.

@@

	This enables RISscatter to softly bridge, generalize, and outperform \gls{bc}, \gls{sr}, \gls{ris}; boil down to any of those three technologies under some specific reflection states; or evolve in a mixed state.
	We show RIScatter is subject to a fundamental tradeoff that balances its information modulation and channel reconfiguration capabilities in a flexible and mutualistic manner.

@@

	An application scenario is considered where a multi-antenna \gls{ap} serves a single-antenna user surrounded by multiple RIScatter nodes, and the achievable primary-total-backscatter rate region is characterized by optimizing input distribution at the nodes, active beamforming at the \gls{ap}, and backscatter decision regions at the user.

@@

	Simulation results demonstrate RIScatter can exploit the additional propagation paths to provide a smooth transition between backscatter modulation and passive beamforming via smart input distribution design, which includes \gls{bc}, \gls{sr}, and \gls{ris} as special cases.

@@

	On the other hand, \gls{ris} is a promising technology that evolves wireless propagation environment using numerous passive reflecting elements (e.g., scattering antenna or programmable metamaterial) with adjustable amplitudes and/or phases \cite{Wu2021b}.
	The scattered signals contain no additional information, but adds constructively or destructively with the direct component to enhance desired signal or suppress interference.
	Compared with backscatter nodes of \gls{bc}/\gls{sr}, \gls{ris} elements employ deterministic reflection pattern priorly known at transmitter and receiver.
	This motivated the use of fixed reflection coefficients during each channel block to improve communication, sensing, and power performances \cite{Wu2018,Zhang2019a,Lin2022,Liu2022,Feng2022,Zhao2022}.
	The concept of dynamic \gls{ris}, namely choosing independent reflection coefficients over different time slots within channel block, was first considered for resource blocks of \gls{ofdm} systems, then extended to power and information phases of \gls{wpcn} \cite{Wu2021,Wu2021d,Hua2022a}.
	Dynamic \gls{ris} provides artificial channel diversity and flexible resource allocation, but misses the opportunity to encode its own message.
	From an information-theoretic perspective, \cite{Karasik2020} reported using \gls{ris} as an auxiliary passive beamforming device to maximize the \gls{snr} is generally rate-suboptimal for finite input constellations.
	Instead, joint transmitter-\gls{ris} encoding achieves the capacity of \gls{ris}-aided channel, and layered encoding with \gls{sic} decoding (i.e., \gls{sic}-based \gls{sr}) can outperform pure passive beamforming at high \gls{snr}.
	It inspired \cite{Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hu2021b,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} to employ \gls{ris} also as an information source to combine passive beamforming and backscatter modulation in the overall reflection pattern.
	In particular, \emph{symbol level precoding} maps backscatter symbols to \gls{ris} coefficient sets optimized for detection \cite{Liu2019d,Bereyhi2020}, \emph{overlay modulation} superposes information-bearing symbols over a common auxiliary matrix \cite{Xu2020b,Zhang2021d,Hu2021b,Hua2022}, \emph{spatial modulation} switches between reflection coefficient sets that maximize \gls{snr} at different receive antennas \cite{Basar2020,Ma2020a,Yuan2021}, and \emph{index modulation} divides \gls{ris} into reflection elements for passive beamforming and information elements for on-off modulation \cite{Hu2021a}.
	However, those \gls{ris}-empowered \gls{bc}/\gls{sr} designs involve advanced hardware architecture, high optimization complexity, and additional control overhead.

@@

	\emph{Fifth,} we provide numerical results to demonstrate the benefits of RIScatter and proposed algorithms.
	It is concluded that 1) RIScatter nodes provide flexible transmit-assist tradeoff using adaptive input distribution design; 2) when primary link is absolutely prioritized, the input probability of each node is \num{1} at one state and \num{0} at the others, which coincides with discretized \gls{ris}; 3) when backscatter link is absolutely prioritized, RIScatter nodes can exploit adaptive channel coding to achieve higher instantaneous rate than uncoded equiprobable inputs of \gls{bc}/\gls{sr}; 4) the proposed \gls{kkt} input distribution design converges to stationary points, whose average performance is very close to global optimal; 5) the proposed \gls{pgd} beamforming provides larger rate region than \gls{mrt}-based schemes; 6) the decision region schemes adapted to input distribution provide higher backscatter rates than the non-adaptive \gls{ml} detector; 7) cooperation between passive nodes in terms of joint encoding can further boost the total backscatter rate; 8) the low complexity of input distribution design and the marginal effect of increasing reflection states motivates the use of low-order scatter nodes to replace high-resolution \gls{ris}.

@@

	\begin{subsection}{BackCom/SR: Backscatter Modulation}
		\gls{bc} and \gls{sr} commonly employ antenna-based scatterers as passive information sources.
		By \emph{frequently switching} between the available states, each backscatter node encodes self message over scattered signal and introduces backscatter uncertainty at the receiver.
		For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by
		% \cite{Thomas2012a}
		\begin{equation}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		\end{equation}
		where $\alpha \in \mathbb{I}$ is the amplitude scattering ratio at the direction of interest.
	\end{subsection}

@@

\begin{section}{Simulation Results}
	In this section, we provide numerical results to evaluate the proposed input distribution, active beamforming, and backscatter decision design for the RIScatter system in Fig. \ref{fi:uniscatter_network}.
	We assume the AP-user distance is \qty{10}{\meter}, $K=2$ RIScatter nodes are randomly dropped within a disk of radius \qty{2}{\meter} centered at the user, the \gls{ap} is with $Q=4$ transmit antennas and an average power budget $P=\qty{36}{\dBm}$, the nodes employs \num{4}-\gls{qam} with amplitude scattering ratio $\alpha=0.5$, the user is with average noise power $\sigma_v^2=\qty{-40}{\dBm}$, and the symbol period ratio is $N=20$.
	For all channels involved, we consider a distance-dependent path loss model
	\begin{equation}
		L(d) = L_0 \biggl(\frac{d_0}{d}\biggr)^\gamma,
	\end{equation}
	together with a Ricean fading model
	\begin{equation}
		\boldsymbol{H} = \sqrt{\frac{\kappa}{1+\kappa}} \bar{\boldsymbol{H}} + \sqrt{\frac{1}{1+\kappa}} \tilde{\boldsymbol{H}}
	\end{equation}
	where $d$ is the transmission distance, $L_0=-\qty{30}{\dB}$ is the reference path loss at distance $d_0=\qty{1}{\meter}$, $\kappa$ is the Ricean $K$-factor, $\bar{\boldsymbol{H}}$ is the deterministic line-of-sight component with entries of unit magnitude, and $\tilde{\boldsymbol{H}}$ is the Rayleigh fading component with entries in standard \gls{iid} \gls{cscg} distribution.
	We choose $\gamma_{\mathrm{D}}=2.6$, $\gamma_{\mathrm{F}}=2.4$, $\gamma_{\mathrm{B}}=2$, and $\kappa_{\mathrm{D}}=\kappa_{\mathrm{F}}=\kappa_{\mathrm{B}}=5$ for direct, forward and backward links.
	The finite threshold domain $\mathcal{T}$ comes from $b$-bit uniform discretization over the critical interval defined by the confidence bounds of edge hypotheses (i.e., lower bound of $\mathcal{H}_1$ and upper bound of $\mathcal{H}_L$) with confidence level $1-\varepsilon$, where $b=7$ and $\varepsilon=\num{1e-3}$.
	All achievable rate curves/regions are averaged over \num{3000} realizations, and the parameters remain fixed unless otherwise specified.

>>>>

	% \gls{bc} switches load impedance to manipulate the magnitude, phase, and/or frequency responses and modulate information onto the scattered signal.
	\gls{bc} switches device load impedance to manipulate the magnitude, phase, and/or frequency responses for information modulation on the scattered carrier.

@@

	% \gls{sr} introduces a passive scatterer to ride secondary information over the primary radio from an active transmitter, and employs a co-located receiver to decode both links.
	% \gls{sr} introduces a passive scatterer to ride secondary information over the primary radio from an active transmitter and employs a co-located receiver to decode both links.
	\gls{sr} allows a passive scattering node to ride additional information over the primary radio emitted by an active transmitter, and employs a co-located receiver to decode both sources.
	% ride secondary information over the primary radio from an active transmitter and employs a co-located receiver to decode both links.
	\gls{ris} programs an antenna/metamaterial array to adjust the scattering phase shifts for signal propagation control in specific directions.
	% and control signal strength in specific directions.

@@

	RIScatter is a new paradigm for future wireless networks and consists of multiple passive scattering nodes, whose reflection states can be controlled to partially engineer the wireless channel and partially modulate information onto the scattered wave.
	This contrasts with \gls{bc}/\gls{sr} (resp. \gls{ris}) where states are exclusively a function of information symbols (resp. \gls{csi}).
	The key principle in RIScatter is to render the probability distribution of reflection states as a joint function of the \gls{csi} and backscatter information source, allowing universal hardware design and heterogeneous traffic control.

@@

	This enables RISscatter to softly bridge, generalize, and outperform \gls{bc}, \gls{sr}, \gls{ris}; boil down to any of those under specific reflection states; or evolve in a mixed setup.
	We reveal RIScatter experiences a fundamental tradeoff that balances information modulation and channel reconfiguration capabilities in a flexible and mutualistic manner.

@@

	An application scenario is also considered where a multi-antenna \gls{ap} serves a single-antenna user surrounded by multiple RIScatter nodes, and the achievable primary-total-backscatter rate region is characterized by optimizing input distribution at the nodes, active beamforming at the \gls{ap}, and backscatter decision regions at the user.

@@

	Simulation results demonstrate RIScatter can exploit the additional propagation paths to provide a smooth transition between backscatter modulation and passive beamforming through smart input distribution design, including \gls{bc}, \gls{sr}, \gls{ris} as special cases.

@@

	On the other hand, \gls{ris} is a promising technology that evolves wireless propagation environment using numerous passive reflecting elements (e.g., scattering antenna or programmable metamaterial) with adjustable amplitudes and/or phases \cite{Wu2021b}.
	The scattered signals contain no additional information, but adds constructively or destructively with the direct component to enhance desired signal or suppress interference.
	Compared with scattering nodes of \gls{bc}/\gls{sr}, \gls{ris} elements employ deterministic reflection pattern priorly known at transmitter and receiver.
	This motivated the use of fixed reflection coefficients during each channel block to improve communication, sensing, and power performances \cite{Wu2018,Zhang2019a,Lin2022,Liu2022,Feng2022,Zhao2022}.
	The concept of dynamic \gls{ris}, namely choosing independent reflection coefficients over different time slots within channel block, was first considered for resource blocks of \gls{ofdm} systems, then extended to power and information phases of \gls{wpcn} \cite{Wu2021,Wu2021d,Hua2022a}.
	Dynamic \gls{ris} provides artificial channel diversity and flexible resource allocation, but misses the opportunity to encode its own message.
	From an information-theoretic perspective, \cite{Karasik2020} reported using \gls{ris} as an auxiliary passive beamforming device to maximize the \gls{snr} is generally rate-suboptimal for finite input constellations.
	Instead, joint transmitter-\gls{ris} encoding achieves the capacity of \gls{ris}-aided channel, and layered encoding with \gls{sic} decoding (i.e., \gls{sic}-based \gls{sr}) can outperform pure passive beamforming at high \gls{snr}.
	It inspired \cite{Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hu2021b,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} to employ \gls{ris} also as an information source to combine passive beamforming and backscatter modulation in the overall reflection pattern.
	In particular, \emph{symbol level precoding} maps backscatter symbols to \gls{ris} coefficient sets optimized for detection \cite{Liu2019d,Bereyhi2020}, \emph{overlay modulation} superposes information-bearing symbols over a common auxiliary matrix \cite{Xu2020b,Zhang2021d,Hu2021b,Hua2022}, \emph{spatial modulation} switches between reflection coefficient sets that maximize \gls{snr} at different receive antennas \cite{Basar2020,Ma2020a,Yuan2021}, and \emph{index modulation} divides \gls{ris} into reflection elements for passive beamforming and information elements for on-off modulation \cite{Hu2021a}.
	However, those \gls{ris}-empowered \gls{bc}/\gls{sr} designs involve advanced hardware architecture, high optimization complexity, and additional control overhead.

@@

	\emph{Fifth,} we provide numerical results to demonstrate the benefits of RIScatter and proposed algorithms.
	It is concluded that 1) RIScatter nodes provide flexible transmit-assist tradeoff using adaptive input distribution design; 2) when primary link is absolutely prioritized, the input probability of each node is \num{1} at one state and \num{0} at the others, which coincides with discretized \gls{ris}; 3) when backscatter link is absolutely prioritized, RIScatter nodes can exploit adaptive channel coding to achieve higher instantaneous rate than uncoded equiprobable inputs of \gls{bc}/\gls{sr}; 4) the proposed \gls{kkt} input distribution design converges to stationary points, whose average performance is very close to global optimal; 5) the proposed \gls{pgd} beamforming provides larger rate region than \gls{mrt}-based schemes; 6) the decision region schemes adapted to input distribution provide higher backscatter rates than the non-adaptive \gls{ml} detector; 7) cooperation between passive nodes in terms of joint encoding can further boost the total backscatter rate; 8) the low complexity of input distribution design and the marginal effect of increasing reflection states motivates the use of low-order scattering nodes to replace high-resolution \gls{ris}.

@@

	\begin{subsection}{BackCom/SR: Backscatter Modulation}
		\gls{bc} and \gls{sr} commonly employ antenna-based scatterers as passive information sources.
		By \emph{frequently switching} between the available states, each scattering node encodes self message over scattered signal and introduces backscatter uncertainty at the receiver.
		For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by
		% \cite{Thomas2012a}
		\begin{equation}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		\end{equation}
		where $\alpha \in \mathbb{I}$ is the amplitude scattering ratio at the direction of interest.
	\end{subsection}

@@

\begin{section}{Simulation Results}
	In this section, we provide numerical results to evaluate the proposed input distribution, active beamforming, and backscatter decision design for the RIScatter system in Fig. \ref{fi:uniscatter_network}.
	We assume the AP-user distance is \qty{10}{\meter}, $K=2$ RIScatter nodes are randomly dropped within a disk of radius \qty{2}{\meter} centered at the user, the \gls{ap} is with $Q=4$ transmit antennas and an average power budget $P=\qty{36}{\dBm}$, the nodes employs \num{4}-\gls{qam} with amplitude scattering ratio $\alpha=0.5$, the user is with average noise power $\sigma_v^2=\qty{-40}{\dBm}$, and the symbol period ratio is $N=20$.
	For all channels involved, we consider a distance-dependent path loss model
	\begin{equation}
		L(d) = L_0 \biggl(\frac{d_0}{d}\biggr)^\gamma
	\end{equation}
	together with a Ricean fading model
	\begin{equation}
		\boldsymbol{H} = \sqrt{\frac{\kappa}{1+\kappa}} \bar{\boldsymbol{H}} + \sqrt{\frac{1}{1+\kappa}} \tilde{\boldsymbol{H}},
	\end{equation}
	where $d$ is the transmission distance, $L_0=-\qty{30}{\dB}$ is the reference path loss at distance $d_0=\qty{1}{\meter}$, $\kappa$ is the Ricean $K$-factor, $\bar{\boldsymbol{H}}$ is the deterministic line-of-sight component with entries of unit magnitude, and $\tilde{\boldsymbol{H}}$ is the Rayleigh fading component with entries in standard \gls{iid} \gls{cscg} distribution.
	We choose $\gamma_{\mathrm{D}}=2.6$, $\gamma_{\mathrm{F}}=2.4$, $\gamma_{\mathrm{B}}=2$, and $\kappa_{\mathrm{D}}=\kappa_{\mathrm{F}}=\kappa_{\mathrm{B}}=5$ for direct, forward and backward links.
	The finite threshold domain $\mathcal{T}$ comes from $b$-bit uniform discretization over the critical interval defined by the confidence bounds of edge hypotheses (i.e., lower bound of $\mathcal{H}_1$ and upper bound of $\mathcal{H}_L$) with confidence level $1-\varepsilon$, where $b=7$ and $\varepsilon=\num{1e-3}$.
	All achievable rate curves/regions are averaged over \num{3000} realizations, and the parameters remain fixed unless otherwise specified.
